Comparing Cross Validation Methods for Penalized Cox Regression

### 2018-03-21 cross-validation.tex
Revert "change figures"
This reverts commit 3e81db9eeadf3261a6054f0af237f4a617f40c44.

\usepackage[margin=1in]{geometry}

@@

\section*{Abstract}
Due to the semi-parametric nature of Cox Regression, conducting cross validation for Cox models has always been a challenge. While cross validation is a commonly used approach for selecting tuning parameters in penalized regression, little research has been done to study cross validation methods for penalized Cox regression. We propose two new cross-validation methods for Cox Regression, and compare them to traditional information criteria as well as a cross-validated partial likelihood approach originally proposed by Verweij et al. Our simulation studies show that, in general, cross-validation tends to be conservative (i.e., select smaller models than the ideal choice of tuning parameters) for penalized Cox regression models. However, our proposed approach of cross-validating the linear predictors generally offers the best balance of stability and performance. We also illustrate these approaches on data from studies of gene expression and progression-free survival in cancer patients.

@@

\begin{figure}
    \centering
		\includegraphics[height= 4cm ]{./figures/cv.png}	
    \caption{K-fold Cross Validation}
\end{figure}

@@

\begin{figure}
    \centering
		\includegraphics[height= 4cm ]{./figures/02_1.png}
		\includegraphics[height= 4cm ]{./figures/02_2.png}
    \caption{Grouped vs Ungrouped Cross Validated Likelihood}
\end{figure}

@@

   \begin{figure}[h]
    \centering
    \begin{minipage}[b]{0.45\textwidth}
      \centering
		  \includegraphics[height= 3.9cm ]{./figures/03.png}
      \caption{CV Linear Predictors}
     \end{minipage}
     \begin{minipage}[b]{0.45\textwidth}
      \centering
		  \includegraphics[height= 4.1cm ]{./figures/04_2.png}
      \caption{CV Deviance Residuals}
      \end{minipage}	
   \end{figure}

@@

@@

   \begin{figure}[h]
   \centering
    \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/05.jpeg}
      \caption{ }
     \end{minipage}
     \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/06.jpeg}
      \caption{ }
      \end{minipage}	
   \end{figure}

@@

  \begin{table}[h]
    \centering
			\begin{tabular}{ c | c | c | c}
				CV Methods & $\lambda$ (SD) & MSE Ratio \\
				\hline \hline
				Optimal  & 0.0606 ( 0.0019 )&  / & \\ 
				Grouped  &  0.0895  ( 0.0043 )& 1.598  \\
				Ungrouped & 0.0754 ( 0.0060 )& 1.225 \\
				Linear Pred & 0.0759 ( 0.0058 )& 1.234 \\
				M Residual &0.1283 ( 0.0047 ) & 2.959
			\end{tabular}
			\caption{Simulation Results in High Dimensional Setting}
	\end{table}

@@

\begin{figure}
    \centering
		\includegraphics[height= 8cm ]{./figures/stability2.png}
    \caption{Comparing Stability}
\end{figure}

    \subsection {Comparisons with Information Criteria}
	%need to change the figures to have fewer lines
    \par We also compared cross-validation with information criteria AIC and BIC. Simulation studies were conducted in lower dimension and higher dimension respectively. As is illustrated in Figure 7, in lower dimension, AIC tends to perform more liberal and BIC tends to perform more conservative than the optimal $\lambda$. While the cross-validation methods are also conservative, most of them are less conservative than BIC. In terms of MSE, cross-validation methods yield smaller MSE than the information criteria. In higher dimension, both BIC and AIC tend to be more liberal than the optimal $\lambda$ and yield larger MSE.

   \begin{figure}[h]
    \centering
    \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/aic_01.jpeg}
      \caption{Comparison to Information Criteria in Lower Dimension}
     \end{minipage}
     \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/aic_02.jpeg}
      \caption{Comparison to Information Criteria in Higher Dimension}
      \end{minipage}	
   \end{figure}

 \subsection {Issues with Deviance Residual}

@@

\begin{table}[h]
\centering
  \begin{tabular}{ c | c | c | c}
				CV Methods & $\lambda$ & $log(\lambda)$ &AUC \\
				\hline \hline
				Grouped  &  0.117 & -2.14 &   0.595\\
				Ungrouped &  0.112 & -2.20 &  0.597\\
				Linear Pred &  0.085 & -2.46 &     0.607\\
				M Residual & 0.171 & -1.76 & 0.529 
	\end{tabular}
  \caption{Tuning Parameter Selected for Ovarian Cancer Data}
\end{table}

@@

   \begin{figure}[h]
   \centering
    \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/07.jpeg}
      \caption{CVE for Ovarian Cancer Data}
     \end{minipage}
     \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/08.jpeg}
      \caption{CVE for Lung Cancer Data}
      \end{minipage}	
   \end{figure}

@@

			\begin{table}[h]
			\centering
			\begin{tabular}{ c | c | c | c}
				CV Methods & $\lambda$ & $log(\lambda)$ & AUC \\
				\hline \hline
				Grouped  &   0.119 & -2.12 &  0.608 \\
				Ungrouped &  0.119 & -2.12 &  0.608 \\
				Linear Pred &  0.090 & -2.41 &  0.633\\
				M Residual & 0.203 & -1.59 &  0.564
			\end{tabular}
			\caption{Tuning Parameter Selected for Lung Cancer Data}
			\end{table}

>>>>

\usepackage[bmargin=0.75in, tmargin =0.75in,lmargin = 0.5in,rmargin = 0.5in]{geometry}

@@

\section*{Abstract}
While cross validation is commonly used for selecting tuning parameters in penalized regression, little research has been done to study cross validation methods for penalized Cox regression. Due to its semi-parametric nature, conducting cross validation for Cox models has always been a challenge. We propose two new cross-validation methods for Cox Regression, and compare them to traditional information criteria as well as a cross-validated partial likelihood approach originally proposed by Verweij et al. Our simulation studies show that, in general, cross-validation tends to be conservative (i.e., select smaller models than the ideal choice of tuning parameters) for penalized Cox regression models. However, our proposed approach of cross-validating the linear predictors generally offers the best balance of stability and performance. We also illustrate these approaches on data from studies of gene expression and progression-free survival in cancer patients.

@@

%\begin{figure}
 %   \centering
	%	\includegraphics[height= 4cm ]{./figures/cv.png}	
    %\caption{K-fold Cross Validation}
%\end{figure}

@@

@@

  % \begin{figure}[h]
    %\centering
    %\begin{minipage}[b]{0.45\textwidth}
      %\centering
	%	  \includegraphics[height= 3.9cm ]{./figures/03.png}
      %\caption{CV Linear Predictors}
     %\end{minipage}
     %\begin{minipage}[b]{0.45\textwidth}
      %\centering
%		  \includegraphics[height= 4.1cm ]{./figures/04_2.png}
    %  \caption{CV Deviance Residuals}
     % \end{minipage}	
  % \end{figure}

@@

\begin{figure}
    \centering
		\includegraphics[height= 9cm ]{./figures/figure_1.png}
		%\includegraphics[height= 4cm ]{./figures/02_2.png}
    \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
\end{figure}

@@

\begin{figure}[h]
    \centering
		\includegraphics[height= 7cm ]{./figures/figure_2.png}
    \caption{Simulation Studies Comparing 4 CV methods in terms of $\lambda$ and MSE ratio. n = 200, p = 1000, magnitude of non-$\beta$s varies from 0.5 to 1.5}
\end{figure}

@@

\begin{table}[h]
\small
\centering
\caption{Simulations Comparing all four Cross-Validation methods, Varying n and p}
\begin{tabular}{llllllllllllllll}
\hline\\[-0.75em]
       & \multicolumn{3}{l}{n = 100, p = 20} &  & \multicolumn{3}{l}{n = 100, p = 100} &  & \multicolumn{3}{l}{n = 100, p = 1000} &  & \multicolumn{3}{l}{n = 500, p = 10000} \\
       & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} \\ \cline{2-4} \cline{6-8} \cline{10-12} \cline{14-16} 
\hline\\[-0.75em]
       &     &        &MSE       &  &   &       & MSE     &    &    &        & MSE      &  &     &        & MSE      \\
       & $\lambda$(SD)     & &Ratio       &  & 	$\lambda$(SD)         & & Ratio     & &	$\lambda$(SD)         & & Ratio      & &	 $\lambda$(SD)         & & Ratio      
\\ 
\cline{2-2} \cline{4-4} \cline{6-6} \cline{8-8} \cline{10-10} \cline{12-12} \cline{14-14} \cline{16-16} 
\\[-0.75em]
Optimal &0.016 (0.009)&&/&&0.042 (0.003)&&/&&0.056 (0.006)&&/&&0.026 (0.001)&&/\\
Grouped &0.035 (0.008)&&2.643&&0.074 (0.009)&&2.236&&0.177 (0.054)&&2.278&&0.053 (0.003)&&77.541\\
Ungrouped &0.030 (0.009)&&2.404&&0.064 (0.010)&&1.801&&0.115 (0.044)&&1.549&&0.041 (0.003)&&56.105\\
LinearPred &0.033 (0.006)&&2.423&&0.066 (0.009)&&1.850&&0.122 (0.039)&&1.673&&0.041 (0.003)&&59.693\\
DevResid &0.099 (0.009)&&13.941&&0.108 (0.009)&&4.110&&0.175 (0.052)&&2.273&&0.075 (0.003)&&76.882\\
\hline
\end{tabular}
\end{table}

@@

\begin{figure}
    \centering
		\includegraphics[height= 7cm ]{./figures/figure_3.png}
    \caption{Comparing Stability}
\end{figure}

    \subsection {Results of Leave-One-Out Cross Validation}
	%need to change the figures to have fewer lines
    \par We also compared cross-validation with information criteria AIC and BIC. Simulation studies were conducted in lower dimension and higher dimension respectively. As is illustrated in Figure 7, in lower dimension, AIC tends to perform more liberal and BIC tends to perform more conservative than the optimal $\lambda$. While the cross-validation methods are also conservative, most of them are less conservative than BIC. In terms of MSE, cross-validation methods yield smaller MSE than the information criteria. In higher dimension, both BIC and AIC tend to be more liberal than the optimal $\lambda$ and yield larger MSE.

\begin{figure}
    \centering
		\includegraphics[height= 8cm ]{./figures/figure_4.png}
    \caption{LOOCV}
\end{figure}

  \subsection {Simulations on Deviance Residuals}

\begin{table}[h]
	\small
	\caption{Average of Selected $\lambda$ with Different Baseline Estimation for Deviance Residuals}
	\centering
	\begin{tabular}{lllllllll}
	\hline
	\\[-0.75em]
    & Optimal & Linear Pred & KP & KP & Breslow & Breslow & Exact & Exp\\ 
    &  & & & weighted & & weighted & & \\ \cline{2 - 9} 
\textbf{True Baseline: $Exponential(1)$} &&&&&&&&\\
n = 100, p = 10, non-zero $\beta$: 2 & 0.0461 & 0.0665 & 0.1986 & 0.1970 & 0.1908 & 0.1975 & 0.0540 & 0.0539 \\
n = 100, p = 100, non-zero $\beta$: 20 &0.0555 & 0.1841 & 0.2359 & 0.2201 & 0.2146 & 0.2244 & 0.0805 & 0.0985 \\
n = 100, p = 1000, non-zero $\beta$: 20 & 0.0268 & 0.0505 & 0.0948 & 0.0798 & 0.0867 & 0.0809 & 0.0288 & 0.0270 \\
\textbf{True Baseline:} $Weibull$ &&&&&&&& \\
 n = 100, p = 10, non-zero $\beta$: 2 & 0.0476 & 0.0710 & 0.2004 & 0.1969 & 0.1907 & 0.1907 & N/A & 0.2330 \\ 
\hline
\end{tabular}

\end{table}

@@

\begin{table}[]
\centering
\caption{Application to Real Data Sets}
\label{my-label}
\begin{tabular}{llllll}
\hline
\\[-0.75em]
 & Ovarian & & & Shedden&  \\
& $\lambda$ & log($\lambda$) &  &$\lambda$  &  log($\lambda$)   \\ \cline{2-3} \cline{5-6} 
Grouped &0.1226 & -2.0985 && 0.1215 & -2.1079\\
Ungrouped & 0.0803 & -2.5222 && 0.0983 & -2.3197\\
Linear Pred &0.0828 & -2.4919 && 0.1013 & -2.2894\\
Dev Resid & 0.0992 & -2.3104 && 0.1110 & -2.1987\\ \hline
\end{tabular}
\end{table}

\begin{figure}
    \centering
		\includegraphics[height= 8cm ]{./figures/figure_5.png}
    \caption{Comparing Cross-Validated Error for Different Methods on Real Data Sets}
\end{figure}

@@

@@
